---
title: "jags_survival"
author: "DLN"
date: "June 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###############################################################################################
######7. Estimation of survival from capture-recapture data using the Cormack-Jolly-Seber (CJS) model
###############################################################################################

```{r}
library(stringr)
library(R2WinBUGS)
library(R2jags)

library(tidyr)
library(dplyr)

setwd("C:/Users/MTHallworth/Documents/DESIREE/demography chickadees/demography analysis/JAGS")
```
#Data 
```{r}
## Upload Data
## Chickadee Encounters
       
ch<-read.csv("C:/Users/MTHallworth/Documents/DESIREE/demography chickadees/demography analysis/data management/capturehistory_2017-03-01.csv",
             colClasses = "character") %>%
    mutate(bandnum=bandNumberQuery) %>%
    mutate(ch=capture_history) %>%
    select(site, bandnum,ch, years_seen)



## Year Data
year<-read.csv("C:/Users/MTHallworth/Documents/DESIREE/demography chickadees/demography analysis/csv/yearvalues_12feb2017.csv") %>%
  select(year, peakcherry_juldate, average_wintertemp)

## Body Condition
bodycondition<-read.csv("C:/Users/MTHallworth/Documents/DESIREE/demography chickadees/demography analysis/bodycondition/CACH_bodycondition_2017-03-01.csv") %>%
  mutate(site=site.x) %>%
   select(bandnum, age, sex, condition)

## Imperv
#imperv<-read.csv("C:/Users/dnarango/Desktop/GIS SPATIAL LAYERS - IMPERV/Imperv_buffer_100_2017-03-01.csv") %>%
#  mutate(site=id) %>%  
#  select(site, per_imperv)
    


## Plants
exotic<-read.csv("C:/Users/MTHallworth/Documents/DESIREE/demography chickadees/veg/plantorigin_IV- 2017-03-23.csv") %>%
  select(site, exotic_tree,exotic_shrub) %>%
  mutate(exotic_type=ifelse(exotic_tree>100,"exotic","native"))


```
Data Manipulation
```{r message=FALSE}


## Add Data Together
survival<-ch %>%
    #left_join(imperv, by="site") %>%
    left_join(bodycondition, by="bandnum") %>%
    left_join(exotic, by="site")
          
## Adds the mean where there are NAs
#survival$per_imperv[is.na(survival$per_imperv)]<-mean(na.omit(survival$per_imperv))


## Include only adults with known sex
adults<-survival %>%
        filter(age!="HY")%>%
        filter(age!="L")%>%
        filter(age!="U")%>%
        filter(age!="ASY?") %>%
        #filter(sex!="M?") %>%
        #filter(sex!="UNK") %>%
        #filter(sex!="U")
        filter(exotic_tree!="NA")

adults<-na.omit(adults)


```

Construct Matrix
```{r}
### need a matrix with x: capture year and y: individual

#[,1] [,2] [,3] [,4] [,5] [,6]
#[1,]    1    0    0    0    0    0
#[2,]    1    0    0    0    0    0
#[3,]    1    1    1    1    1    1
#[4,]    1    1    0    0    0    0
#[5,]    1    0    0    0    0    0
#[6,]    1    1    0    0    1    0

ch<-as.matrix(str_split_fixed(adults$ch, "", 15))

class(ch) <- "numeric" 
matrix<-ch

## also need vector with marking occassion (year marked)
get.first <- function(x) min(which(x!=0)) #make function
VECTOR <- apply(matrix, 1, get.first)
```

######################################################################################
#### BUGS

A base model
```{r}
bugs.dir=("C:/Program Files (x86)/WinBUGS14")

# Specify model in BUGS language
sink("cjs-c-c.bug")
cat("                          ##starts the bugs code
    model {
      
# Priors and constraints
      phi ~ dunif(0, 1)            ### Uniform priors  (could use: phi[i,t] <- mean.phi for mean survival)
      p ~ dunif(0, 1)             ### Uniform priors
      
      
      # Likelihood 
      for (i in 1:nind){
        # Define latent state at first capture
        z[i,f[i]] <- 1
        for (t in (f[i]+1):n.occasions){
          
          # State process
          z[i,t] ~ dbern(mu1[i,t])
          mu1[i,t] <- phi * z[i,t-1]		# Depends on whether ind was alive the time before
          
          # Observation process
          y[i,t] ~ dbern(mu2[i,t])
          mu2[i,t] <- p * z[i,t]			# Depends on the whether ind alive that time period
        } #t
      } #i
    }
    ",fill = TRUE)
sink()

# Bundle data
bugs.data <- list(y = matrix, f = VECTOR, nind = nrow(matrix), n.occasions = ncol(matrix))

# Function to create a matrix of initial values for latent state z
ch.init <- function(ch, f){
  for (i in 1:dim(ch)[1]){ch[i,1:f[i]] <- NA}
  return(ch)
}

# Initial values
inits <- function(){list(z = ch.init(matrix, VECTOR), phi = runif(1, 0, 1), p = runif(1, 0, 1))}

# Parameters monitored
parameters <- c("phi", "p")

# MCMC settings
ni <- 1000
nt <- 6
nb <- 500
nc <- 3

# Call WinBUGS from R (BRT 1 min)

directory<-"C:/Users/MTHallworth/Documents/DESIREE/demography chickadees/demography analysis/BUGS code"

cjs.c.c <- bugs(bugs.data, inits, parameters, "cjs-c-c.bug", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, debug = TRUE, bugs.directory = bugs.dir, working.directory = getwd())

# Summarize posteriors
print(cjs.c.c, digits = 3)
```


# To run in JAGS
# In JAGS we have to give good initial values for the latent state z. At all occasions when an individual was 
# observed, its state is z = 1 for sure. In addition, if an individual was not observed at an occasion, 
#but was alive for sure, because it was observed before and thereafter (i.e. has a capture history of e.g. {101} 
#or {10001}), then we know that the individual was alive at all of these occasions, and thus z = 1. 
#Therefore, we should provide initial values of z = 1 at these positions as well. The following function 
#provides such initial values from the observed capture histories:
```{r}
known.state.cjs <- function(ch){
  state <- ch
  for (i in 1:dim(ch)[1]){
    n1 <- min(which(ch[i,]==1))
    n2 <- max(which(ch[i,]==1))
    state[i,n1:n2] <- 1
    state[i,n1] <- NA
  }
  state[state==0] <- NA
  return(state)
}

JAGS_inits <- function(){list(phi = runif(1, 0, 1), p = runif(1, 0, 1), z = known.state.cjs(matrix))}

cjs.c.cJ <- jags(bugs.data, JAGS_inits, parameters, "cjs-c-c.bug", n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)

# Summarize posteriors
print(cjs.c.cJ, digits = 3)
```
